Contract Orchestration for OCaml \\ Progress Report

### 2021-06-30 reports/2021-07-01/COOC-report.tex
sync

\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{hyperref}

\title{Contract Orchestration for OCaml \\
  Progress Report 
}
\date{\today}
\author{Prof.\ Dr.\ Peter Thiemann, University of Freiburg, Germany}
\begin{document}
\maketitle{}

@@

@@

\paragraph{October 2020 - June 2021}
\begin{itemize}
\item Tamara Bernhardt worked her master's project on distributed assertion checking for Tezos smart contracts. The assertion is specified in its own syntax in a separate file from the contract implementation. This scenario requires a separate compiler for assertions in Michelson. The generated code is either combined with the contract implementation or stored separately on the blockchain. Tamara worked on grammar, lexer/parser, transformer and the compiler for the assertion language. 
\begin{itemize}
\item Defining grammar for the assertion language. Although the grammar is related to Michelson, it was designed to be as general as possible so that it can be adapted to different smart contract languages. There are two types of grammars: prefix and infix.
\item Implementing and extensively testing the lexer/parser using OCamllex and Menhir for the prefix style version.
\item Writing a backend transformer for a specific target. The backend interface is builted as a virtual library for which multiple implementations can be provided.
\item Implementing the compiler that generates the target code according to an on-chain orchestration scheme between the assertion and the parent contract.
\end{itemize}
At the end, she write the development document and then wrap it up in her thesis.
\end{itemize}

\paragraph{November 2020 - June 2021}
\begin{itemize}
\item We recruited Julian Veigel to do his master's thesis with us. He was our student in the blockchain course. For his thesis, he will be developing a checking assertion for Ethereum. He and Tamara will collaborate on some parts of the assertion project, e.g. they will collaborate on formulating this assertion and implementing it in generators and/or exchange ideas. He completed his thesis in mid-June. It deals with different approaches on how to implement such a validation process. The basic idea is to distribute the validation process to several nodes in the network, all testing different property cases. The goal is to reduce the cost while preserving basic properties such as decentralization and trust. All the approaches are discussed in detail and the most promising one is shown in an implementation. The contribution of his work includes:
\begin{itemize}
\item exploring Ethereum infrastructures, especially the recent release of Ethereum 2.0.
\item presenting different approaches and discuss their strengths and weaknesses.
\item laying some mathematical groundwork to build a qualitative.
\item an implementation with its components for one of the presented approaches.
\end{itemize}
\end{itemize}

\paragraph{April 2021}
\begin{itemize}
\item Leqaa Ahmed, a bachelor student in computer science, starts her bachelor thesis in my group. 
\end{itemize}

Ha and I have been running the course “Blockchain and Cryptocurrencies”. This course builds on the book “Bitcoin and Cryptocurrency Technologies” by Narayanan and others. It covers seven weeks worth of material from the book; two weeks on Ethereum, Solidity, and its security issues; and two weeks on Tezos, with the bulk of material on Michelson.
\paragraph{May 2020 - September 2020}
\begin{itemize}
\item We run the course for the first time. We designed the exercise to help students confirm their understanding of the basic concepts of cryptocurrency and blockchain and also focus on training students in blockchain programming. Students are required to implement some basic blockchain systems in Python, write smart contracts in Solidity and Michelson. In terms of Michelson, we guided the students to understand how the languages work and study its strong type system. For practice, we use some online tools that support coding and testing of smart contracts in Solidity, e.g. Remix ethereum and Michelson, e.g. Try michelson. Finally, we have included the final exam with an open time for 48 hours. The final exam covers the main concepts and programming tasks. The grading of the exam was completed and submitted at the end of October. At the end of the course, we received some feedback from the students. Some of them found the course interesting and some might consider doing research on Blockchain with us in the future.
\end{itemize}

\paragraph{April 2021 - September 2021}
\begin{itemize} 
\item We are running the course for the second time. The basic content of the lecture and exercises has remained essentially the same, but has been adjusted according to feedback we have received from former students. We will be working on the new exam for this year, which will be in September.
\end{itemize}

Ha and I have worked on the safe-interaction with Tezos blockchain.
\paragraph{August 2020  - June 2021}
\begin{itemize} 
\item Tezos provides low-level interfaces, which work with loosely structured data in JSON format that is inconvenient to handle and offers few guarantees. Ha and I proposed Contract modules that provide a statically checked interface to interact with contracts on the Tezos blockchain. A module specification provides all types as well as information about potential failure conditions of the contract. The specification is checked against the contract implementation using symbolic execution. An OCaml module is generated that contains a function for each entrypoint of the contract. The types of these functions fully describe the interface including the failure conditions and guarantee type-safe and sometimes fail-safe invocation of the contract on the blockchain.

Based on the result of the research, we have submited a short paper on towards contract modules for the Tezos blockchain to a workshop on formal Methods for Blockchains.

@@

As a result, we have submited a research paper to a conference on programming Languages and systems. The   
Contributions of the paper is as follows:
\begin{itemize}
\item A typed API for originating and invoking contracts as well as querying the
state of the blockchain.
\item An operational semantics for functional programs running alongside smart
contracts in a blockchain.
\item Established various properties of the combined system with proofs in upcoming techreport.
\item An implementation of a low-level OCaml-API to the Tezos blockchain, which corresponds to the operational semantics. 
\end{itemize}

\end{itemize}

@@

\begin{itemize}
\item  to be written.
\end{itemize}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{hyperref}
\usepackage{authblk}

\title{Contract Orchestration for OCaml \\
  Progress Report 
}
\author{Thi Thu Ha Doan}
\author{Peter Thiemann}
\affil{University of Freiburg, Germany}
\begin{document}
\maketitle{}

@@

\subsection{Personnel}
\label{sec:personnel}

There is one full-time employee working on the project for the whole time period: Dr. Thi Thu Ha Doan (postdoctoral researcher).

Student workers (part time as explained below): Tamara Bernhardt, Joshua Allgeier

@@

\paragraph{October 2020 - June 2021}
\begin{itemize}
\item Tamara Bernhardt worked her master's thesis (thesis work is generally unpaid, but supervised by Dr.\ Doan) on distributed  assertion checking for Tezos smart contracts. The assertion is specified in its own syntax in a separate file from the contract implementation. This scenario requires a separate compiler for assertions.  The generated code is either combined with the contract implementation or stored separately on the blockchain. Tamara worked on grammar, lexer/parser, transformer, and the compiler for the assertion language. 
\begin{itemize}
\item Defining the grammar for the assertion language. Although the underlying type structure is inspired by Michelson, it was designed to be as general as possible so that it can be adapted to different smart contract languages. As Tamara did not have  experience with parsing, she created a parser for a Scheme-style prefix syntax using Menhir. Another student was scheduled  to implement an alternative grammar inspired by OCaml syntax (as a term project), but he did not finish the work.
\item Implementing and extensively testing the lexer/parser using OCamllex and Menhir for the prefix style version.
\item Writing a backend transformer for a specific target. The backend interface is built as a virtual library for which multiple  implementations can be provided.
\item Implementing the compiler that generates the target code according to an on-chain orchestration scheme between the assertion and the parent contract.
\end{itemize}
At the end, Tamara wrote the development document and then wrapped it up in her thesis.
\end{itemize}

\paragraph{November 2020 - June 2021}
\begin{itemize}
\item We supervised the master's thesis of Julian Veigel. He developed a counterpart of the assertion checker for Ethereum. He  and Tamara collaborated on some parts of the assertion project, e.g., on assertion language and general analysis of the problem. Julian completed his thesis in mid-June. It deals with different approaches on how to implement such a validation process. The basic idea is to distribute the validation process to several nodes in the network, all testing different property cases. The goal is to reduce the cost while preserving basic properties such as decentralization and trust. All the approaches are discussed in detail and the most promising one is shown in an implementation. The contribution of his work includes:
\begin{itemize}
\item exploring Ethereum infrastructures, especially the recent release of Ethereum 2.0.
\item presenting different approaches and discuss their strengths and weaknesses.
\item laying some mathematical groundwork to build a qualitative.
\item an implementation with its components for one of the presented approaches.
\end{itemize}
\end{itemize}

Julian's work is not directly related to Tezos and to the project's goals. Nevertheless, we gained a lot of understanding of the problem due to his contributions and due to the additional perspective provided by considering Ethereum.

\paragraph{April 2021}
\begin{itemize}
\item Leqaa Ahmed, a bachelor student in computer science, starts her bachelor thesis in my group. Her task is to further develop and strengthen the typing properties of our Tezos-API. She is mainly supervised by Dr. Doan.
\end{itemize}

Dr. Doan (mainly) and I have been running the course “Blockchain and Cryptocurrencies”. This course builds on the book “Bitcoin and Cryptocurrency Technologies” by Narayanan and others. Our implementation covers seven weeks worth of material from the book; two weeks on Ethereum, Solidity, and its security issues; and two weeks on Tezos, with the bulk of material on Michelson. 
\paragraph{May 2020 - September 2020}
\begin{itemize}
\item First instance of the course with 25 students. We created the lecture materials consisting of slides and videos.  We designed numerous exercise tasks to help students confirm their understanding of the basic concepts of cryptocurrency and blockchain and also focus on training students in blockchain programming. Students are required to implement some basic blockchain systems in  Python. The they write smart contracts in Solidity and Michelson. In terms of Michelson, we guided the students to understand how the languages work and study its strong type system. For practice, we use some online tools that support coding and testing of smart contracts in Solidity (Remix ethereum) and Michelson (Try michelson). Finally, we conducted the final exam in the form of a take home exam during a time window of 48 hours. The final exam covers the main concepts and several programming tasks. The grading of the exam was completed and submitted at the end of October. 
\end{itemize}

\paragraph{April 2021 - September 2021}
\begin{itemize} 
\item Second instance of the course with 39 students. The basic content of the lecture and exercises has remained essentially the same, but has  been adjusted according to feedback we have received from former students. Dr.\ Doan is conducting the weekly exercise sessions while lectures are mainly provided with recordings from the first instance. We plan to create one new lecture to be delivered in July. The exam will be in September in the same format as in the first instance. 
\end{itemize}

Doan and Thiemann have worked on the safe interaction with contracts on the Tezos blockchain.
\paragraph{August 2020  - June 2021}
\begin{itemize} 
\item Tezos provides low-level interfaces, which work with loosely structured data in JSON format that is inconvenient to handle and offers few guarantees. We proposed Contract modules that provide a statically checked interface to interact with contracts on the Tezos blockchain. A module specification provides all types as well as information about potential failure conditions of the contract. The specification is checked against the contract implementation using symbolic execution. An OCaml module is generated that contains a function for each entrypoint of the contract. The types of these functions fully describe the interface including the failure conditions and guarantee type-safe and sometimes fail-safe invocation of the contract on the blockchain.

Based on the result of the research, we submitted a short paper titled ``Towards Contract Modules for the Tezos Blockchain''  to the workshop on Formal Methods for Blockchains (FMBC 2021).

@@

As a result, we submitted a research paper to the Asian Symposium on Programming Languages and Systems (APLAS 2021). The contributions of the paper are as follows:
\begin{itemize}
\item A typed API for originating and invoking contracts as well as querying the
state of the blockchain.
\item An operational semantics for functional programs running alongside smart
contracts in a blockchain.
\item Established various properties of the combined system with proofs in upcoming techreport.
\item An implementation of a low-level OCaml-API to the Tezos blockchain, which corresponds to the operational semantics. 
\end{itemize}
\end{itemize}

@@

To support this paper, we implemented Michelsym, a prototype of a symbolic interpreter for Michelson. 
\begin{itemize}
\item  to be written.
\end{itemize}
